---
title: "Metabarcoding Alpha Diversity Analysis"
author: "Emily Bean"
date: "April 8, 2019"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warnings = FALSE)
# we don't want to display code (echo), warnings (warnings), or the chunk (include) in the output
```

```{r}
## Reading in packages and data
if(!require(prettydoc)) install.packages("prettydoc")
library(prettydoc)
if(!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)
if(!require(tidyverse)) install.packages("tidyverse")
library(tidyverse)
if(!require(tidyverse)) install.packages("emmeans")
library(emmeans)
if(!require(qpcR)) install.packages("qpcR")
library(qpcR)
# 2017 ITS
alphadiv2017ITS <- read.table(file = "//petalibrary.arcc.uwyo.edu/homes/lvandiep/SoilEcologyLab/Students/Bean/HTS/alphaITS2017_NOBLOCK4.txt", header = TRUE, sep = "\t") %>% 
  dplyr::select(-c(ID, se.chao1, se.ACE)) %>% 
  gather(key = "Param", value = "value",
         Observed, Chao1, ACE, Shannon, Simpson, InvSimpson, Fisher)
# 2018 16S
alphadiv201816S <- read.table(file = "//petalibrary.arcc.uwyo.edu/homes/lvandiep/SoilEcologyLab/Students/Bean/HTS/alpha_16S2018NOBLOCK4.txt", header = TRUE, sep = "\t") %>% 
  mutate(
         soil_type = case_when(
           soil_type %in% "all" ~ "combined",
           soil_type %in% "rhiz" ~ "rhizo",
           soil_type %in% "bulk" ~ "bulk"
         )) %>% 
  dplyr::select(-c(ID, se.chao1, se.ACE)) %>% 
    gather(key = "Param", value = "value",
         Observed, Chao1, ACE, Shannon, Simpson, InvSimpson, Fisher)
# 2017 16S
alphadiv201716S <- read.table(file = "//petalibrary.arcc.uwyo.edu/homes/lvandiep/SoilEcologyLab/Students/Bean/HTS/alphadiv_16S2017_NOBLOCK4.txt", header = TRUE, sep = "\t") %>% 
  mutate(
         soil_type = case_when(
           type %in% "combined_bulk_rhizo" ~ "combined",
           type %in% "bulk" ~ "bulk",
           type %in% "rhizo" ~ "rhizo"
         ),
         type = NULL)  %>% 
  dplyr::select(-c(ID, se.chao1, se.ACE)) %>% 
    gather(key = "Param", value = "value",
         Observed, Chao1, ACE, Shannon, Simpson, InvSimpson, Fisher)

# combine and gather to vertical
all16S <- rbind(alphadiv201716S, alphadiv201816S) %>% 
  gather(key = "Param", value = "value",
         Observed, Chao1, ACE, Shannon, Simpson, InvSimpson, Fisher)

```

```{r}
# function to check assumptions of linear regression
myLinReg <- function(par) {
  p1 <- hist(all$value[all$Param == par], 
             main = paste("Hist", par, sep = ": "),
             xlab = as.character(par))
  
  p2 <- qqnorm(all$value[all$Param == par], 
               main = paste("QQPlot", par, sep = ": "))
  p3 <- qqline(all$value[all$Param == par], col = "red")
  
  p4 <- boxplot(all$value[all$Param == par] ~ 
                  all$GrazeTime[all$Param == par], data = all,
                main = paste("Boxplot vs GrazeTime", par, sep = ": "))
  p5 <- boxplot(all$value[all$Param == par] ~ 
                  all$Treatment[all$Param == par], data = all,
                main = paste("Boxplot vs Treatment", par, sep = ": "))
  
  return(list(p1, p2, p3, p4, p5))
}
```

####Alpha diversity analysis

1. Check assumptions of linear regression and normalize data as much as possible
2. GLM structure similar to univariate data: raw numbers instead of % difference from PRE
3. Plot any parameters that the GLM indicates are significant (p < 0.05)
4. 16S data is analyzed (1) bulk only, (2) rhizospheric only, (3) rhizospheric and bulk combined. The 16S data had more reads in the bulk soil than rhizospheric so rareficiation 

*Overall, only 16S rhizospheric samples from 2018 have any significance, and only with one alpha diversity metric at one time*


```{#r}
# Checking assumptions of linear regression - don't print
pars <- unique(all$Param)
yrs <- unique(all$year)
stype <- unique(all$soil_type)
seq <- unique(all$seq)

graphics.off()
for(j in 1:length(yrs)){
  for(n in 1:length(seq)) {
    for(m in 1:length(stype)) {
      for(i in 1:length(pars)){
       
        
          par(mfrow = c(2,2)) 
          print(myLinReg(par = pars[i]))
          dev.off
        
      }
    }
  }
}
  
# the data shows significant kurtosis, which is not able to be easily transformed
```

##### ITS 2017: Only the Simpson index is significant at HDG between 1WK and 4WK

```{r}

# create empty data frames to fill in the loop
posthoc1 <- data.frame()
posthoc2 <- data.frame()
modelFit <- data.frame()

# GLM loop 

pars <- unique(alphadiv2017ITS$Param)
      for(i in 1:length(pars)){
        
        
        
        mod <- glm(value ~ GrazeTime * Treatment,
                   data = filter(alphadiv2017ITS, Param == pars[i]),
                   family = gaussian(link = "identity"))
        mf <- data.frame(Param = as.character(pars[i]),
                         deviance = mod$deviance,
                         null.deviance = mod$null.deviance,
                         diff = mod$null.deviance - mod$deviance,
                         df.null = mod$df.null,
                         df.dev = mod$df.residual,
                         rmse = round(RMSE(mod), 4))
        modelFit <- rbind(mf, modelFit)
        e <- as.data.frame(
          emmeans(mod, pairwise ~ Treatment | GrazeTime, type = "response"))
        df <- e %>% dplyr::select(c(contrasts.contrast, contrasts.GrazeTime,
                                    contrasts.SE, contrasts.z.ratio, 
                                    contrasts.p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc1 <- rbind(df, posthoc1)
        e2 <- emmeans(mod, pairwise ~ GrazeTime | Treatment, type = "response")
        e2 <- as.data.frame(e2$contrasts)
        df1 <- e2 %>% dplyr::select(c(contrast, Treatment,
                                      SE, z.ratio, 
                                      p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc2 <- rbind(df1, posthoc2)
        
        
      }
    
# clean up the model outputs for further analysis
colnames(posthoc1) <- c("contrast", "contTS.Trt", "SE", "z.ratio", 
                        "p.value", "Parameter")
posthoc1$sign <- ifelse(posthoc1$p.value <= 0.05,
                        "significant",
                        "not_significant")
colnames(posthoc2) <- c("contrast", "contTS.Trt", "SE", "z.ratio",
                        "p.value", "Parameter")
posthoc2$sign <- ifelse(posthoc2$p.value <= 0.05,
                        "significant",
                        "not_significant")
# what is significant?
sig.TSITS <- posthoc1[which(posthoc1$sign %in% "significant"),]
sig.TrtITS <- posthoc2[which(posthoc2$sign %in% "significant"),]

outallITS <- rbind(posthoc1, posthoc2)
```
##### 16S 2018 bulk & rhizospheric combined- no significance

```{r}
a16S2017 <- alphadiv201716S[alphadiv201716S$soil_type %in% "combined",]

# create empty data frames to fill in the loop
posthoc1 <- data.frame()
posthoc2 <- data.frame()
modelFit <- data.frame()

# GLM loop 

pars <- unique(a16S2017$Param)
      for(i in 1:length(pars)){
        
        
        
        mod <- glm(value ~ GrazeTime * Treatment,
                   data = filter(a16S2017, Param == pars[i]),
                   family = gaussian(link = "identity"))
        mf <- data.frame(Param = as.character(pars[i]),
                         deviance = mod$deviance,
                         null.deviance = mod$null.deviance,
                         diff = mod$null.deviance - mod$deviance,
                         df.null = mod$df.null,
                         df.dev = mod$df.residual,
                         rmse = round(RMSE(mod), 4))
        modelFit <- rbind(mf, modelFit)
        e <- as.data.frame(
          emmeans(mod, pairwise ~ Treatment | GrazeTime, type = "response"))
        df <- e %>% dplyr::select(c(contrasts.contrast, contrasts.GrazeTime,
                                    contrasts.SE, contrasts.z.ratio, 
                                    contrasts.p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc1 <- rbind(df, posthoc1)
        e2 <- emmeans(mod, pairwise ~ GrazeTime | Treatment, type = "response")
        e2 <- as.data.frame(e2$contrasts)
        df1 <- e2 %>% dplyr::select(c(contrast, Treatment,
                                      SE, z.ratio, 
                                      p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc2 <- rbind(df1, posthoc2)
        
        
      }
    
# clean up the model outputs for further analysis
colnames(posthoc1) <- c("contrast", "contTS.Trt", "SE", "z.ratio", 
                        "p.value", "Parameter")
posthoc1$sign <- ifelse(posthoc1$p.value <= 0.05,
                        "significant",
                        "not_significant")
colnames(posthoc2) <- c("contrast", "contTS.Trt", "SE", "z.ratio",
                        "p.value", "Parameter")
posthoc2$sign <- ifelse(posthoc2$p.value <= 0.05,
                        "significant",
                        "not_significant")
# what is significant?
sig.TS16S17 <- posthoc1[which(posthoc1$sign %in% "significant"),]
sig.Trt16S17 <- posthoc2[which(posthoc2$sign %in% "significant"),]

outall16S17 <- rbind(posthoc1, posthoc2)
```
##### 16S 2017 - bulk - no significance

```{r}
b16S2017 <- alphadiv201716S[alphadiv201716S$soil_type %in% "bulk",]

# create empty data frames to fill in the loop
posthoc1 <- data.frame()
posthoc2 <- data.frame()
modelFit <- data.frame()

# GLM loop 

pars <- unique(b16S2017$Param)
      for(i in 1:length(pars)){
        
        
        
        mod <- glm(value ~ GrazeTime * Treatment,
                   data = filter(b16S2017, Param == pars[i]),
                   family = gaussian(link = "identity"))
        mf <- data.frame(Param = as.character(pars[i]),
                         deviance = mod$deviance,
                         null.deviance = mod$null.deviance,
                         diff = mod$null.deviance - mod$deviance,
                         df.null = mod$df.null,
                         df.dev = mod$df.residual,
                         rmse = round(RMSE(mod), 4))
        modelFit <- rbind(mf, modelFit)
        e <- as.data.frame(
          emmeans(mod, pairwise ~ Treatment | GrazeTime, type = "response"))
        df <- e %>% dplyr::select(c(contrasts.contrast, contrasts.GrazeTime,
                                    contrasts.SE, contrasts.z.ratio, 
                                    contrasts.p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc1 <- rbind(df, posthoc1)
        e2 <- emmeans(mod, pairwise ~ GrazeTime | Treatment, type = "response")
        e2 <- as.data.frame(e2$contrasts)
        df1 <- e2 %>% dplyr::select(c(contrast, Treatment,
                                      SE, z.ratio, 
                                      p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc2 <- rbind(df1, posthoc2)
        
        
      }
    
# clean up the model outputs for further analysis
colnames(posthoc1) <- c("contrast", "contTS.Trt", "SE", "z.ratio", 
                        "p.value", "Parameter")
posthoc1$sign <- ifelse(posthoc1$p.value <= 0.05,
                        "significant",
                        "not_significant")
colnames(posthoc2) <- c("contrast", "contTS.Trt", "SE", "z.ratio",
                        "p.value", "Parameter")
posthoc2$sign <- ifelse(posthoc2$p.value <= 0.05,
                        "significant",
                        "not_significant")
# what is significant?
sig.TS16S17B <- posthoc1[which(posthoc1$sign %in% "significant"),]
sig.Trt16S17B <- posthoc2[which(posthoc2$sign %in% "significant"),]

outall16S17B <- rbind(posthoc1, posthoc2)
```

##### 16S 2017 Rhizospheric - no significance
```{r}
r16S2017 <- alphadiv201716S[alphadiv201716S$soil_type %in% "rhizo",]

# create empty data frames to fill in the loop
posthoc1 <- data.frame()
posthoc2 <- data.frame()
modelFit <- data.frame()

# GLM loop 

pars <- unique(r16S2017$Param)
      for(i in 1:length(pars)){
        
        
        
        mod <- glm(value ~ GrazeTime * Treatment,
                   data = filter(r16S2017, Param == pars[i]),
                   family = gaussian(link = "identity"))
        mf <- data.frame(Param = as.character(pars[i]),
                         deviance = mod$deviance,
                         null.deviance = mod$null.deviance,
                         diff = mod$null.deviance - mod$deviance,
                         df.null = mod$df.null,
                         df.dev = mod$df.residual,
                         rmse = round(RMSE(mod), 4))
        modelFit <- rbind(mf, modelFit)
        e <- as.data.frame(
          emmeans(mod, pairwise ~ Treatment | GrazeTime, type = "response"))
        df <- e %>% dplyr::select(c(contrasts.contrast, contrasts.GrazeTime,
                                    contrasts.SE, contrasts.z.ratio, 
                                    contrasts.p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc1 <- rbind(df, posthoc1)
        e2 <- emmeans(mod, pairwise ~ GrazeTime | Treatment, type = "response")
        e2 <- as.data.frame(e2$contrasts)
        df1 <- e2 %>% dplyr::select(c(contrast, Treatment,
                                      SE, z.ratio, 
                                      p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc2 <- rbind(df1, posthoc2)
        
        
      }
    
# clean up the model outputs for further analysis
colnames(posthoc1) <- c("contrast", "contTS.Trt", "SE", "z.ratio", 
                        "p.value", "Parameter")
posthoc1$sign <- ifelse(posthoc1$p.value <= 0.05,
                        "significant",
                        "not_significant")
colnames(posthoc2) <- c("contrast", "contTS.Trt", "SE", "z.ratio",
                        "p.value", "Parameter")
posthoc2$sign <- ifelse(posthoc2$p.value <= 0.05,
                        "significant",
                        "not_significant")
# what is significant?
sig.TS16S17R <- posthoc1[which(posthoc1$sign %in% "significant"),]
sig.Trt16S17R <- posthoc2[which(posthoc2$sign %in% "significant"),]

outall16S17R <- rbind(posthoc1, posthoc2)
```

#####16S 2018 bulk and rhizospheric - no significance

```{r}
a16S2018 <- alphadiv201816S[alphadiv201816S$soil_type %in% "combined",]

# create empty data frames to fill in the loop
posthoc1 <- data.frame()
posthoc2 <- data.frame()
modelFit <- data.frame()

# GLM loop 

pars <- unique(a16S2018$Param)
      for(i in 1:length(pars)){
        
        
        
        mod <- glm(value ~ GrazeTime * Treatment,
                   data = filter(a16S2018, Param == pars[i]),
                   family = gaussian(link = "identity"))
        mf <- data.frame(Param = as.character(pars[i]),
                         deviance = mod$deviance,
                         null.deviance = mod$null.deviance,
                         diff = mod$null.deviance - mod$deviance,
                         df.null = mod$df.null,
                         df.dev = mod$df.residual,
                         rmse = round(RMSE(mod), 4))
        modelFit <- rbind(mf, modelFit)
        e <- as.data.frame(
          emmeans(mod, pairwise ~ Treatment | GrazeTime, type = "response"))
        df <- e %>% dplyr::select(c(contrasts.contrast, contrasts.GrazeTime,
                                    contrasts.SE, contrasts.z.ratio, 
                                    contrasts.p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc1 <- rbind(df, posthoc1)
        e2 <- emmeans(mod, pairwise ~ GrazeTime | Treatment, type = "response")
        e2 <- as.data.frame(e2$contrasts)
        df1 <- e2 %>% dplyr::select(c(contrast, Treatment,
                                      SE, z.ratio, 
                                      p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc2 <- rbind(df1, posthoc2)
        
        
      }
    
# clean up the model outputs for further analysis
colnames(posthoc1) <- c("contrast", "contTS.Trt", "SE", "z.ratio", 
                        "p.value", "Parameter")
posthoc1$sign <- ifelse(posthoc1$p.value <= 0.05,
                        "significant",
                        "not_significant")
colnames(posthoc2) <- c("contrast", "contTS.Trt", "SE", "z.ratio",
                        "p.value", "Parameter")
posthoc2$sign <- ifelse(posthoc2$p.value <= 0.05,
                        "significant",
                        "not_significant")
# what is significant?
sig.TS16S18 <- posthoc1[which(posthoc1$sign %in% "significant"),]
sig.Trt16S18 <- posthoc2[which(posthoc2$sign %in% "significant"),]

outall16S18 <- rbind(posthoc1, posthoc2)
```

##### 16S 2018 - bulk - no significance

```{r}
b16S2018 <- alphadiv201816S[alphadiv201816S$soil_type %in% "bulk",]

# create empty data frames to fill in the loop
posthoc1 <- data.frame()
posthoc2 <- data.frame()
modelFit <- data.frame()

# GLM loop 

pars <- unique(b16S2018$Param)
      for(i in 1:length(pars)){
        
        
        
        mod <- glm(value ~ GrazeTime * Treatment,
                   data = filter(b16S2018, Param == pars[i]),
                   family = gaussian(link = "identity"))
        mf <- data.frame(Param = as.character(pars[i]),
                         deviance = mod$deviance,
                         null.deviance = mod$null.deviance,
                         diff = mod$null.deviance - mod$deviance,
                         df.null = mod$df.null,
                         df.dev = mod$df.residual,
                         rmse = round(RMSE(mod), 4))
        modelFit <- rbind(mf, modelFit)
        e <- as.data.frame(
          emmeans(mod, pairwise ~ Treatment | GrazeTime, type = "response"))
        df <- e %>% dplyr::select(c(contrasts.contrast, contrasts.GrazeTime,
                                    contrasts.SE, contrasts.z.ratio, 
                                    contrasts.p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc1 <- rbind(df, posthoc1)
        e2 <- emmeans(mod, pairwise ~ GrazeTime | Treatment, type = "response")
        e2 <- as.data.frame(e2$contrasts)
        df1 <- e2 %>% dplyr::select(c(contrast, Treatment,
                                      SE, z.ratio, 
                                      p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc2 <- rbind(df1, posthoc2)
        
        
      }
    
# clean up the model outputs for further analysis
colnames(posthoc1) <- c("contrast", "contTS.Trt", "SE", "z.ratio", 
                        "p.value", "Parameter")
posthoc1$sign <- ifelse(posthoc1$p.value <= 0.05,
                        "significant",
                        "not_significant")
colnames(posthoc2) <- c("contrast", "contTS.Trt", "SE", "z.ratio",
                        "p.value", "Parameter")
posthoc2$sign <- ifelse(posthoc2$p.value <= 0.05,
                        "significant",
                        "not_significant")
# what is significant?
sig.TS16S18B <- posthoc1[which(posthoc1$sign %in% "significant"),]
sig.Trt16S18B <- posthoc2[which(posthoc2$sign %in% "significant"),]

outall16S18B <- rbind(posthoc1, posthoc2)
```

##### 16S 2018 Rhizospheric 
Simpson is significant at NG from PRE-1WK
Simpson & InvSimpson significant from HI-NO at 1WK
```{r}
r16S2018 <- alphadiv201816S[alphadiv201816S$soil_type %in% "rhizo",]

# create empty data frames to fill in the loop
posthoc1 <- data.frame()
posthoc2 <- data.frame()
modelFit <- data.frame()

# GLM loop 

pars <- unique(r16S2018$Param)
      for(i in 1:length(pars)){
        
        
        
        mod <- glm(value ~ GrazeTime * Treatment,
                   data = filter(r16S2018, Param == pars[i]),
                   family = gaussian(link = "identity"))
        mf <- data.frame(Param = as.character(pars[i]),
                         deviance = mod$deviance,
                         null.deviance = mod$null.deviance,
                         diff = mod$null.deviance - mod$deviance,
                         df.null = mod$df.null,
                         df.dev = mod$df.residual,
                         rmse = round(RMSE(mod), 4))
        modelFit <- rbind(mf, modelFit)
        e <- as.data.frame(
          emmeans(mod, pairwise ~ Treatment | GrazeTime, type = "response"))
        df <- e %>% dplyr::select(c(contrasts.contrast, contrasts.GrazeTime,
                                    contrasts.SE, contrasts.z.ratio, 
                                    contrasts.p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc1 <- rbind(df, posthoc1)
        e2 <- emmeans(mod, pairwise ~ GrazeTime | Treatment, type = "response")
        e2 <- as.data.frame(e2$contrasts)
        df1 <- e2 %>% dplyr::select(c(contrast, Treatment,
                                      SE, z.ratio, 
                                      p.value)) %>% 
          mutate(Param = as.character(pars[i]))
        posthoc2 <- rbind(df1, posthoc2)
        
        
      }
    
# clean up the model outputs for further analysis
colnames(posthoc1) <- c("contrast", "contTS.Trt", "SE", "z.ratio", 
                        "p.value", "Parameter")
posthoc1$sign <- ifelse(posthoc1$p.value <= 0.05,
                        "significant",
                        "not_significant")
colnames(posthoc2) <- c("contrast", "contTS.Trt", "SE", "z.ratio",
                        "p.value", "Parameter")
posthoc2$sign <- ifelse(posthoc2$p.value <= 0.05,
                        "significant",
                        "not_significant")
# what is significant?
sig.TS16S18R <- posthoc1[which(posthoc1$sign %in% "significant"),]
sig.Trt16S18R <- posthoc2[which(posthoc2$sign %in% "significant"),]

outall16S18R <- rbind(posthoc1, posthoc2)
```