---
title: "One-Way ANOVAs for C, N, MBC Data 2017"
output: html_notebook
---

Emily Bean  
ebean@uwyo.edu  

***

 Below are a series of one-way ANOVAs completed on the following data:  
 1.Microbial Biomass C  
 2.Microbial Biomass N  
 3.Microbial Biomass C:N Ratio  
 3.Non-purgeable Organic Carbon (NPOC)  
 4.Dissolved Organic Nitrogen (DON)  
 5.Mineral Nitrogen (NO3-N + NH4-N)  
 6.Extracellular Enzymes 
 7.Microbial Efficiency (enzyme activity per unit microbial biomass)  
 8.Rising Plate Pasture Meter (RPM) - by sample  
 9.Rising Plate Pasture Meter (RPM) - by plot  
 
 ***


One-way ANOVAs are testing the differences in means of three Treatments, at four different GrazeTimes. For each dataset, 4 ANOVAs will be completed (one for each GrazeTime).  

  - GrazeTimes compared to PRE baseline data (each plot average subtracted by that plot's PRE baseline average) - will only have 3 ANOVAs for 24H, 1WK, 4WK
  - Each Plot average (n = 12 per GrazeTime, 48 total)  
  - No averages (n = 60 per GrazeTime, 240 total)  
  
All post-hoc tests are paired t test with a Bonferroni correction  

***

```{r, echo = FALSE}
# Reading in data
require(ggplot2)
require(tidyverse)
require(dplyr)
source("https://raw.githubusercontent.com/EmilyB17/grazing_soil_microbes/master/R/prepData.R")

toc <- prepData(read.table(file = "https://raw.githubusercontent.com/EmilyB17/grazing_soil_microbes/master/data/allCNdata_2017.txt", sep = "\t", header = TRUE, stringsAsFactors = TRUE))
```

**Microbial Biomass Carbon (MBC)**

1. GrazeTimes compared to PRE baseline

No significance

```{r}
# ---- Microbial Biomass C ----
mbc <- toc[, c("GrazeTime", "Treatment", "Plot", "Block", "Sample", "mbc_mgkg")]

# Examine data relative to the PRE baseline
all <- merge(aggregate(mbc_mgkg ~ Plot, data = mbc[mbc$GrazeTime %in% "PRE",], mean),
             aggregate(mbc_mgkg ~ GrazeTime + Plot + Treatment, 
                       data = mbc[!mbc$GrazeTime %in% "PRE",], mean), 
             by = "Plot")
all$delta <- all$mbc_mgkg.y - all$mbc_mgkg.x
all$mbc_mgkg.x <- NULL
all$mbc_mgkg.y <- NULL

# Now the data is averaged by Plot, and "delta" is the difference of the average of that plot from the average of the same plot at the PRE baseline

# plot
ggplot(data = all, aes(x = Treatment, y = delta, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(all$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(delta ~ Treatment, data = filter(all, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary)
# No significance
```
MBC  

2. Each Plot average (n = 12 per GrazeTime, 48 total)

No significance
```{r}
avg <- aggregate(mbc_mgkg ~ Plot + Treatment + GrazeTime, data = mbc, mean)
ggplot(data = avg, aes(x = Treatment, y = mbc_mgkg, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(avg$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(mbc_mgkg ~ Treatment, data = filter(avg, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # No significance
```
MBC  

3. No averages (n = 60 per GrazeTime, 240 total)  

The ANOVA indicated significance at 24H, but the post hoc test did not show significance

```{r}
# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(mbc$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(mbc_mgkg ~ Treatment, data = filter(mbc, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # Significance at 24H

# post-hoc test
with(mbc[mbc$GrazeTime %in% "24H",],
     pairwise.t.test(mbc_mgkg, Treatment, p.adjust.method = "bonferroni", paired = TRUE))
# No significance

```
**Microbial Biomass Nitrogen (MBN)**

1. GrazeTimes compared to PRE baseline data (each plot average subtracted by that plot's PRE baseline average)
```{r}
# ---- Microbial Biomass N ----
mbn <- toc[, c("GrazeTime", "Treatment", "Plot", "Block", "Sample", "mbn_mgkg")]

# Examine data relative to the PRE baseline
all <- merge(aggregate(mbn_mgkg ~ Plot, data = mbn[mbn$GrazeTime %in% "PRE",], mean),
             aggregate(mbn_mgkg ~ GrazeTime + Plot + Treatment, 
                       data = mbn[!mbn$GrazeTime %in% "PRE",], mean), 
             by = "Plot")
all$delta <- all$mbn_mgkg.y - all$mbn_mgkg.x
all$mbn_mgkg.x <- NULL
all$mbn_mgkg.y <- NULL

# Now the data is averaged by Plot, and "delta" is the difference of the average of that plot from the average of the same plot at the PRE baseline

# plot
ggplot(data = all, aes(x = Treatment, y = delta, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(all$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(delta ~ Treatment, data = filter(all, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary)
# No significance
```
MBN  

2. Each Plot average (n = 12 per GrazeTime, 48 total)  

No significance
```{r}

avg <- aggregate(mbn_mgkg ~ Plot + Treatment + GrazeTime, data = mbn, mean)
ggplot(data = avg, aes(x = Treatment, y = mbn_mgkg, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(avg$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(mbn_mgkg ~ Treatment, data = filter(avg, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # No significance

```
MBN  

3. No averages (n = 60 per GrazeTime, 240 total)  

No significance
```{r}
# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(mbn$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(mbn_mgkg ~ Treatment, data = filter(mbn, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # No significance
```
**Microbial Biomass C:N Ratio**

1. GrazeTimes compared to PRE baseline data (each plot average subtracted by that plot's PRE baseline average)

No significance
```{r}
# ---- Microbial Biomass C:N Ratio ----
ratio <- toc$mbc_mgkg / toc$mbn_mgkg
smb <- cbind(toc[,c("GrazeTime", "Treatment", "Block", "Plot", "Sample")], ratio)

# Examine data relative to the PRE baseline
all <- merge(aggregate(ratio ~ Plot, data = smb[smb$GrazeTime %in% "PRE",], mean),
             aggregate(ratio ~ GrazeTime + Plot + Treatment, 
                       data = smb[!smb$GrazeTime %in% "PRE",], mean), 
             by = "Plot")
all$delta <- all$ratio.y - all$ratio.x
all$ratio.x <- NULL
all$ratio.y <- NULL
# Now the data is averaged by Plot, and "delta" is the difference of the average of that plot from the average of the same plot at the PRE baseline

# plot
ggplot(data = all, aes(x = Treatment, y = delta, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(all$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(delta ~ Treatment, data = filter(all, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary)
# No significance
```
Microbial Biomass C:N Ratio  

2.Each Plot average (n = 12 per GrazeTime, 48 total)  

No significance
```{r}
avg <- aggregate(ratio ~ Plot + Treatment + GrazeTime, data = smb, mean)
ggplot(data = avg, aes(x = Treatment, y = ratio, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(avg$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(ratio ~ Treatment, data = filter(avg, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # No significance
```
Microbial Biomass C:N Ratio 

3. No averages (n = 60 per GrazeTime, 240 total)  

No significance

```{r}
# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(smb$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(ratio ~ Treatment, data = filter(smb, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # No significance
```

**Non-purgeable Organic Carbon (NPOC)**  

1. GrazeTimes compared to PRE baseline data (each plot average subtracted by that plot's PRE baseline average)  

No significance
```{r}
# ---- NPOC ----
npoc <- toc[, c("GrazeTime", "Treatment", "Plot", "Block", "Sample", "NPOC_mgkg")]

# Examine data relative to the PRE baseline
all <- merge(aggregate(NPOC_mgkg ~ Plot, data = npoc[npoc$GrazeTime %in% "PRE",], mean),
             aggregate(NPOC_mgkg ~ GrazeTime + Plot + Treatment, 
                       data = npoc[!npoc$GrazeTime %in% "PRE",], mean), 
             by = "Plot")
all$delta <- all$NPOC_mgkg.y - all$NPOC_mgkg.x
all$NPOC_mgkg.x <- NULL
all$NPOC_mgkg.y <- NULL

# Now the data is averaged by Plot, and "delta" is the difference of the average of that plot from the average of the same plot at the PRE baseline

# plot
ggplot(data = all, aes(x = Treatment, y = delta, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(all$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(delta ~ Treatment, data = filter(all, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary)
# No significance
```

NPOC  

2. Each Plot average (n = 12 per GrazeTime, 48 total)  

No significance
```{r}
avg <- aggregate(NPOC_mgkg ~ Plot + Treatment + GrazeTime, data = npoc, mean)
ggplot(data = avg, aes(x = Treatment, y = NPOC_mgkg, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(avg$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(NPOC_mgkg ~ Treatment, data = filter(avg, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # No significance
```

NPOC  

3. No averages (n = 60 per GrazeTime, 240 total)  

The ANOVA indicated significance at 1WK, but the post hoc test did not show significance

```{r}
# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(npoc$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(NPOC_mgkg ~ Treatment, data = filter(npoc, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # Significance at 1WK

# post-hoc test
with(npoc[npoc$GrazeTime %in% "1WK",],
     pairwise.t.test(NPOC_mgkg, Treatment, p.adjust.method = "bonferroni", paired = TRUE))
# No significance

```
**Dissolved Organic Nitrogen (DON)**  

1. GrazeTimes compared to PRE baseline data (each plot average subtracted by that plot's PRE baseline average)   

No significance
```{r}
# ---- DON ----
don <- toc[, c("GrazeTime", "Treatment", "Plot", "Block", "Sample", "DON_mgkg")]

# Examine data relative to the PRE baseline
all <- merge(aggregate(DON_mgkg ~ Plot, data = don[don$GrazeTime %in% "PRE",], mean),
             aggregate(DON_mgkg ~ GrazeTime + Plot + Treatment, 
                       data = don[!don$GrazeTime %in% "PRE",], mean), 
             by = "Plot")
all$delta <- all$DON_mgkg.y - all$DON_mgkg.x
all$DON_mgkg.x <- NULL
all$DON_mgkg.y <- NULL

# Now the data is averaged by Plot, and "delta" is the difference of the average of that plot from the average of the same plot at the PRE baseline

# plot
ggplot(data = all, aes(x = Treatment, y = delta, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(all$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(delta ~ Treatment, data = filter(all, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary)
# No significance
```
DON 

2. Each Plot average (n = 12 per GrazeTime, 48 total)  

No significance

```{r}
avg <- aggregate(DON_mgkg ~ Plot + Treatment + GrazeTime, data = don, mean)
ggplot(data = avg, aes(x = Treatment, y = DON_mgkg, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(avg$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(DON_mgkg ~ Treatment, data = filter(avg, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # No significance
```

DON  

3. No averages (n = 60 per GrazeTime, 240 total)  

No significance

```{r}
# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(don$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(DON_mgkg ~ Treatment, data = filter(don, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # No significance
```

**Mineral Nitrogen (NO3-N + NH4-N)**

1. GrazeTimes compared to PRE baseline data (each plot average subtracted by that plot's PRE baseline average)  

No significance
```{r}
# ---- Mineral N ----
min <- toc[, c("GrazeTime", "Treatment", "Plot", "Block", "Sample", "minN_mgkg")]

# Examine data relative to the PRE baseline
all <- merge(aggregate(minN_mgkg ~ Plot, data = min[min$GrazeTime %in% "PRE",], mean),
             aggregate(minN_mgkg ~ GrazeTime + Plot + Treatment, 
                       data = min[!min$GrazeTime %in% "PRE",], mean), 
             by = "Plot")
all$delta <- all$minN_mgkg.y - all$minN_mgkg.x
all$minN_mgkg.x <- NULL
all$minN_mgkg.y <- NULL

# Now the data is averaged by Plot, and "delta" is the difference of the average of that plot from the average of the same plot at the PRE baseline

# plot
ggplot(data = all, aes(x = Treatment, y = delta, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(all$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(delta ~ Treatment, data = filter(all, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary)
# No significance
```

Mineral N  

2. Each Plot average (n = 12 per GrazeTime, 48 total)

No significance

```{r}
avg <- aggregate(minN_mgkg ~ Plot + Treatment + GrazeTime, data = min, mean)
ggplot(data = avg, aes(x = Treatment, y = minN_mgkg, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(avg$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(minN_mgkg ~ Treatment, data = filter(avg, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # No significance
```

Mineral N  

3. No averages (n = 60 per GrazeTime, 240 total)  

No significance

```{r}
# perform series of one-way ANOVAs, one for each GrazeTime
grz <- unique(min$GrazeTime)
mods<- list()
for(i in 1:length(grz)) {
  mods<- c(mods, list(aov(minN_mgkg ~ Treatment, data = filter(min, GrazeTime == grz[i]))))
}

names(mods)<- grz
lapply(mods, summary) # No significance
```
**Extracellular Enzymes**

1.  GrazeTimes compared to PRE baseline data (each plot average subtracted by that plot's PRE baseline average)  

No significance

```{r}
# ---- Enzymes ----

enz.vert <- prepData(read.table(file = "https://raw.githubusercontent.com/EmilyB17/grazing_soil_microbes/master/data/enzymes_vertical.txt", sep = "\t", header = TRUE, stringsAsFactors = TRUE))

# Examine data relative to the PRE baseline
all <- merge(aggregate(Enzyme_nm_g_hr ~ Plot + Substrate, data = enz.vert[enz.vert$GrazeTime %in% "PRE",], mean),
             aggregate(Enzyme_nm_g_hr ~ GrazeTime + Plot + Treatment + Substrate, 
                       data = enz.vert[!enz.vert$GrazeTime %in% "PRE",], mean), 
             by = c("Plot", "Substrate"))
all$delta <- all$Enzyme_nm_g_hr.y - all$Enzyme_nm_g_hr.x
all$Enzyme_nm_g_hr.x <- NULL
all$Enzyme_nm_g_hr.y <- NULL

# perform one-way ANOVA
sub <- unique(all$Substrate)
grz <- unique(all$GrazeTime)
mods<- list()
nm <- character()
for(j in 1:length(sub)){
  for(i in 1:length(grz)) {
    mods<- c(mods, list(aov(delta ~ Treatment, 
                          data = filter(all, GrazeTime == grz[i] & Substrate == sub[j]))))
    nm <- c(nm, paste(grz[i], sub[j], sep = "_"))
}}
names(mods)<- nm
lapply(mods, summary)
# No significance
```

Extracellular Enzymes

2. Each Plot average (n = 12 per GrazeTime, 48 total)

No significance
```{r}
avg <- aggregate(Enzyme_nm_g_hr ~ Plot + Treatment + GrazeTime + Substrate, 
                 data = enz.vert, mean)
ggplot(data = avg, aes(x = Treatment, y = Enzyme_nm_g_hr, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform repeated-measures ANOVA
sub <- unique(avg$Substrate)
grz <- unique(avg$GrazeTime)
# save the model objects
mods<- list()
nm <- character()
for(j in 1:length(sub)){
  for(i in 1:length(grz)) {
    mods<- c(mods, list(aov(Enzyme_nm_g_hr ~ Treatment, 
                          data = filter(avg, GrazeTime == grz[i] & Substrate == sub[j]))))
    nm <- c(nm, paste(grz[i], sub[j], sep = "_"))
}}
names(mods)<- nm
lapply(mods, summary) 
# No significance
```

Extracellular Enzymes  

3. No averages (n = 60 per GrazeTime, 240 total per substrate)

The ANOVA showed no significance between any unique combination of Subtrate & GrazeTime, between Treatments.

```{r}
# perform repeated-measures ANOVA
sub <- unique(enz.vert$Substrate)
grz <- unique(enz.vert$GrazeTime)
# save the model objects
mods<- list()
nm <- character()
for(j in 1:length(sub)){
  for(i in 1:length(grz)) {
    mods<- c(mods, list(aov(Enzyme_nm_g_hr ~ Treatment, 
                          data = filter(enz.vert, GrazeTime == grz[i] & Substrate == sub[j]))))
    nm <- c(nm, paste(grz[i], sub[j], sep = "_"))
}}
names(mods)<- nm
lapply(mods, summary) 
# No significance
```
**Microbial Efficiency (enzyme activity per unit microbial biomass)**

1. GrazeTimes compared to PRE baseline data (each plot average subtracted by that plot's PRE baseline average)  

The ANOVA showed a significance at 24H NAG, but the post hoc test did not show significance.

```{r}
# ---- Microbial Efficiency ----
# AKA: Enzyme activity per unit microbial biomass 
# Enzyme_nm_g_hr / mbc_mgkg

# Create efficiency (microbial efficiency)
dat <- merge(mbc, enz.vert, by = c("Plot", "GrazeTime", "Block", "Treatment", "Sample"))
dat$efficiency <- dat$Enzyme_nm_g_hr / dat$mbc_mgkg
dat$efficiency[dat$efficiency %in% "Inf"] <- 0

# Examine data relative to the PRE baseline
all <- merge(aggregate(efficiency ~ Plot + Substrate, 
                       data = dat[dat$GrazeTime %in% "PRE",], mean),
             aggregate(efficiency ~ GrazeTime + Plot + Treatment + Substrate, 
                       data = dat[!dat$GrazeTime %in% "PRE",], mean), 
             by = c("Plot", "Substrate"))
all$delta <- all$efficiency.y - all$efficiency.x
all$efficiency.x <- NULL
all$efficiency.y <- NULL

# Now the data is averaged by Plot, and "delta" is the difference of the average of that plot from the average of the same plot at the PRE baseline

# perform one-way ANOVA
sub <- unique(all$Substrate)
grz <- unique(all$GrazeTime)
mods<- list()
nm <- character()
for(j in 1:length(sub)){
  for(i in 1:length(grz)) {
    mods<- c(mods, list(aov(delta ~ Treatment, 
                          data = filter(all, GrazeTime == grz[i] & Substrate == sub[j]))))
    nm <- c(nm, paste(grz[i], sub[j], sep = "_"))
}}
names(mods)<- nm
lapply(mods, summary) # Significance at 24H NAG

# post hoc test for 24H NAG
with(all[all$Substrate %in% "NAG" & all$GrazeTime %in% "24H",], 
     pairwise.t.test(delta, Treatment, p.adjust.method = "bonferroni", paired = TRUE)) # No significance
```

Microbial efficiency 

2. Each Plot average (n = 12 per GrazeTime, 48 total)

Only the NAG substrate at the PRE GrazeTime showed significance, but the post hoc test did not show significance.
```{r}

avg <- aggregate(efficiency ~ Plot + Treatment + GrazeTime + Substrate, 
                 data = dat, mean)
ggplot(data = avg, aes(x = Treatment, y = efficiency, colour = Treatment)) +
  geom_boxplot() +
  facet_wrap(~ GrazeTime)

# perform one-way
sub <- unique(avg$Substrate)
grz <- unique(avg$GrazeTime)
# save the model objects
mods<- list()
nm <- character()
for(j in 1:length(sub)){
  for(i in 1:length(grz)) {
    mods<- c(mods, list(aov(efficiency ~ Treatment, 
                          data = filter(avg, GrazeTime == grz[i] & Substrate == sub[j]))))
    nm <- c(nm, paste(grz[i], sub[j], sep = "_"))
}}
names(mods)<- nm
lapply(mods, summary) # Significance with NAG PRE

# post hoc test
with(avg[avg$GrazeTime %in% "PRE" & avg$Substrate %in% "NAG",],
     pairwise.t.test(efficiency, Treatment, p.adjust.method = "bonferroni", paired = TRUE)) # No significance
```

Microbial Efficiency
MBC  

3. No averages (n = 60 per GrazeTime, 240 total)  

The ANOVA indicated significance at 24H BX, 24H BG, 24H LAP
```{r}
# perform one-way
sub <- unique(dat$Substrate)
grz <- unique(dat$GrazeTime)
# save the model objects
mods<- list()
nm <- character()
for(j in 1:length(sub)){
  for(i in 1:length(grz)) {
    mods<- c(mods, list(aov(efficiency ~ Treatment, 
                          data = filter(dat, GrazeTime == grz[i] & Substrate == sub[j]))))
    nm <- c(nm, paste(grz[i], sub[j], sep = "_"))
}}
names(mods)<- nm
lapply(mods, summary) # Significance with 24H BX, 24H BG, 24H LAP


```
Microbial Efficiency

3. Individual post hoc tests for 24H BX, 24H BG, 24H LAP

```{r}
## 24H BX, 24H BG, 24H LAP

# 24H BX
with(dat[dat$GrazeTime %in% "24H" & avg$Substrate %in% "BX",],
     pairwise.t.test(efficiency, Treatment, p.adjust.method = "bonferroni", paired = TRUE))
```

